\mytitle

### 2015-08-02 MSc_diss.tex
Rewriting literature review

The photochemical formation of the alkyl nitrates starts when the chemistry of the troposphere reaches the regime of $O_3$ production, i.e. when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ concentration. In this case the reaction between of the alkyl peroxy radicals ($RO_2$ where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}, which is controlled by the branching ratio $\alpha_3$ in the following way:
\begin{subequations} \label{eq:peroxy_no0}
\begin{align}
RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{eq:peroxy_no1}\\
RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{eq:peroxy_no2}
\end{align}
\end{subequations}
The former reaction is a crucial since it can serve as a termination step

in the net production of tropospheric ozone, since photolysis of $NO_2$ is the main photochemical source of $O_3$ in the troposphere.

In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of alkyl nitrates via rearrangement of a high-energy intermediate \citep{Bertman1995}. The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.

As it is shown in Figure 2 reactions (8) and (9) form part of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by a radical termination reactions (not shown). However, the reaction of $CH_3O_2$ with NO is a two-channel reaction \citep{Day2003}, and its second branch can terminate the $O_3$ forming cycle earlier by producing an alkyl nitrate, in this case methyl nitrate ($CH_3ONO_2$):
\begin{equation} \label{eq:CH3NO3form}
CH_3O_2 + NO \rightarrow CH_3ONO_2
\end{equation}

Being a reservoir for the short lived NOx and having atmospheric lifetimes ranging from several days to about a month \citep{Reeves2007} alkyl nitrates can be transported over long distances, and then lost through photolysis and reaction with $OH$:
\begin{equation} \label{eq:CH3NO3+hv}
CH_3ONO_2 + hv \rightarrow CH_3O + NO_2
\end{equation}
\begin{equation} \label{eq:CH3NO3+OH}
CH_3ONO_2 + OH \rightarrow HCHO + NO_2
\end{equation}
It means that alkyl nitrates can ‘release’ $NO_2$ back to the atmosphere after a while, and in association with reactions (10) and (11) start a new $O_3$ formation cycle being quite away from the source.

\subsubsection*{Bertman}
Then the alkyl peroxy radical reacts with $NO$, when it can either lose an oxygen atom to form alkoxy radical ($RO\bullet$) or bond with $NO$ to form an alkyl nitrate. The former reaction is a crucial in the net production of tropospheric ozone, since photolysis of $NO_2$ is the main photochemical source of $O_3$ in the troposphere.
\begin{subequations} \label{eq:peroxy_no0}
\begin{align}
ROO\bullet + NO &\xrightarrow{k_3, (1-\alpha_3)} RO\bullet + NO_2 \label{eq:peroxy_no1}\\
ROO\bullet + NO &\xrightarrow{k_3, \alpha_3} RONO_2 \label{eq:peroxy_no2}
\end{align}
\end{subequations}
In other words, a branching ratio $\alpha_3$ of the reactions between alkyl peroxy radicals and $NO$ leads to formation of alkyl nitrates via rearrangement of a high-energy intermediate \citep{Bertman1995}. The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.

\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance = 4cm, auto]
\node[reg] (rh) {$RH$};
\node[rad] (roo) at (4,0){$ROO\bullet$};
\node[add] (oh) at (3,0.3) {$\bullet OH$};
\node[rad] (roor) at (4,-2){$ROOR'$};
\node[add] (r_oo) at (3.7,-1) {$R'OO\bullet$};
\node[reg] (rono2) at (2,3){$RONO_2$};
\node[rad] (ro) at (6,3){$RO\bullet$};
\node[add] (alpha) at ($(roo)!.5!(rono2) + (-1,-0.2)$) {$\alpha_3$};
\node[add] (alpha2) at ($(roo)!.5!(ro) + (1,-0.2)$) {$1-\alpha_3$};
\node[add] (no) at ($(roo)+(0,1.5)$) {$NO$};
\node[reg] (rcho) at (9,3){$RCHO$};
\node[add] (prod) at (0,3){\textit{products}};

\draw[ar] (rh.east) to [out=0,in=180](roo.west);
\draw[ar] (roo.south) to [out=270,in=90](roor.north);
\draw[ar] (roo.north) to [out=90,in=270](rono2.south);
\draw[ar] (roo.north) to [out=90,in=270](ro.south);
\draw[ar] (rono2.west) to [out=180,in=0](prod.east);
\draw[ar] (ro.east) to [out=0,in=180](rcho.west);
\end{tikzpicture}
\caption{Schematic representation of the alkyl nitrate chemistry and their formation through the photooxydation of alkanes. Adapted from \citep{Bertman1995}.}
\label{fig:scheme_bertman}
\end{figure}

Apart from $NO$, alkyl peroxy radicals can react with each other and give peroxides, without formation of nitrates. As the eqs. \eqref{eq:peroxy_no0} outline, whether $ROO\bullet$ will bond with nitrates or with other peroxy radicals depends partially on $NO$ concentration. The reaction between $ROO\bullet$ and $NO$ is more likely when nitrate concentration is high.

The alkyl nitrates are removed from the atmosphere through photolysis and reaction with hydroxyl radical.
\begin{equation} \label{eq:an_photo}
RONO_2 + h\nu \xrightarrow{j_4} RO\bullet + NO_2
\end{equation}
\begin{equation} \label{eq:an_oh}
RONO_2 + OH \xrightarrow{k_5} \mathit{products}
\end{equation}
The relative importance of these two loss reactions depends generally on the net UV intensity and the size of the carbon chain in alkyl nitrates. Photolysis is relatively more important for nitrates with $< C_4$, while nitrates with longer chains ($\geq C_5$) tend to react faster with $OH$. In case of butyl nitrates both mechanisms are of similar significance \citep{Worton2010}.

[Assuming that the only source of alkyl nitrates is the $OH$ reaction with alkane, and that reactions take place in $NO$-rich atmosphere (so no $ROO\bullet$ self-reactions occur), the alkyl nitrate part in this generalised scheme can be reduced to the two equations of formation and loss:
\begin{subequations} \label{eq:an_form_loss0}
\begin{align}
RH &\xrightarrow{\beta k_A} RONO2 \label{eq:an_form_loss1}\\
RONO_2 &\xrightarrow{k_B} \mathit{products} \label{eq:an_form_loss2}
\end{align}
\end{subequations}
where $k_A = k_1[OH]$ and $k_B=j_4 + k_5[OH]$. The factor $\beta=\alpha_1\alpha_3$ takes into account the fraction of $H$ atom abstraction from the particular hydrocarbon (with the reaction rate $k_1$) and the branching ratios leading to the nitrate. The photolytic loss rate of $RONO_2$ is denoted by $j_4$, and $k_5$ is the rate of alkyl nitrate destruction by $OH$.

These kinetic equations lead to the ordinary differential equations which can be integrated in time ($t$) giving a relationship between the alkyl nitrates and their parent alkanes concentrations:
\begin{equation} \label{eq:integral}
\frac{[RONO_2]}{[RH]} = \frac{\beta k_A}{k_B - k_A}\left(1-e^{(k_A - k_B)t}\right)+\frac{[RONO_2]_0}{[RH]_0}e^{(k_A - k_B)t}
\end{equation}
where subscript $0$ denotes the initial concentrations.

A further good approximation is that there is no alkyl nitrate initially (\textit{i.e.} no direct emissions), hence, the second term in eq. (\eqref{eq:integral}) becomes zero. Finally, the evolution of alkyl nitrates in an air mass can be modelled using the following formula.
\begin{equation} \label{eq:model}
\frac{[RONO_2]}{[RH]} = \frac{\beta k_A}{k_B - k_A}\left(1-e^{(k_A - k_B)t}\right)
\end{equation}
It has to be noted that eq. \eqref{eq:model} assumes no mixing with surrounding air. This simplification is reasonable only when mixing affects both the parent alkane and the corresponding alkyl nitrate similarly, keeping the ratios in eqs. \eqref{eq:integral} and \eqref{eq:model} the same \citep{Reeves2007}].

@@

\subsection{Motivation}
lalalaep

@@

mean ANs concentrations: \citep{Reeves2007}, \citep{Roberts1998}

\subsubsection{Modelling studies}

@@

!! from \citep{Finlayson-Pitts2000}
The relationship between the peroxy radical concentration and the ozone photolysis rate constant for these higher NO conditions can be again approximated using steady-state analysis [..]. While OH is recycled in its reactions with CO and CH4 via HO2, it is pernanetly removed at higher NOx concentrations by the reaction of OH with NO2, forming nitric acid:
OH + NO2 = M = HONO2 (113).
This reaction is primarily a daytime reaction because most OH sources are photolytic in nature. As a result, the NO2 reaction with OH competes!!!! with NO2 photolysis (page 290/993)!!
         Rate coeff (OH+NO2 = HNO3)  J(NO2 = NO+O)
Book     2.47e-30          7.00e-3
My model 1.15e-11          4.39e-03
Thus, the reaction with OH is not usually a dominant loss process for NO2, but it is still sufficiently fact to from significant amounts of HNO3 during the day, particularly in polluted regions with relatively large NO2 concentrations.
Nitric acid undergoes both wet and dry deposition rapidly and can be neutralized by ammonia, the major gaseous base found in the atmosphere.

>>>>

The photochemical formation of the alkyl nitrates starts when the chemistry of the troposphere reaches the regime of $O_3$ production, i.e. when reactions of the peroxy radicals with $NO$ start to compete with their self- and cross-reactions more effectively due to an increase in $NO_x$ concentration. In this case the reaction between the alkyl peroxy radical ($RO_2$, where $R$ denotes methyl group, $CH_3$) and $NO$ can reveal its two-channel nature \citep{Day2003}, which is controlled by the branching ratio $\alpha_3$ in the following way \citep{Roberts1990}
\begin{subequations} \label{reac:RONO2ab}
\begin{align}
RO_2 + NO &\xrightarrow{1-\alpha_3} RO + NO_2 \label{eq:peroxy_no1}\\
RO_2 + NO &\xrightarrow{\alpha_3} RONO_2 \label{eq:peroxy_no2}
\end{align}
\end{subequations}
The latter reaction can serve as a termination step of the propagated $O_3$ forming cycle, and as \citep{Roberts1998} estimated this likely begin to occur when the $NO$ concentration increases up to 100-200 ppt.

\subsubsection*{Branching ratio}
The branching ratio $\alpha_3$ is larger when the parent alkane is larger and is a function of temperature and pressure.
\citep{Newland2013}
[The yield of this secondary pathway increases with increasing carbon chain length (Table 1.3)]

\subsubsection*{Yield}
\citep{Newland2013}
[Since the yield of alkyl nitrates from Reaction 1.9 increases with chain length but the mixing ratios of the alkanes tends to decrease with chain length, propyl and butyl nitrates are generally present at the highest concentrations in polluted air (Beyersdorf et al., 2010)]
\citep{Sommariva2008}
[The yield of alkyl nitrate formation is temperature and pressure dependent and increases with the number of carbons in the peroxy radical \citep{Roberts1990}!]

\subsubsection*{Alternative photochemical chemical sources}
\citep{Newland2013}
[Another photochemical pathway has been shown to become important for the production of the lower (methyl) alkyl nitrates in polluted air. Reaction 1.10 shows the reaction of the alkoxy radical, formed in in Reaction 1.9a, with NO2.
\begin{equation}\label{reac:RO+NO2=RONO2}
RO + NO_2 \rightarrow RONO_2
\end{equation}
This photochemical pathway has been shown to become dominant over Reaction 1.9b for the production of methyl nitrate in polluted air masses where NOx mixing ratios exceed 40 ppb (Archibald et al., 2007). This reaction is only likely to be important for methyl nitrate (!) rather than the higher nitrates because of the very small branching ratio ($\alpha_3$) for the production of methyl nitrate via the reaction of the peroxy radical with NO]
\citep{Newland2013}
[in highly polluted environments, the reaction of the alkoxy radical with the NO2 radical (Reaction 1.10) may be important for formation of the C1 – C2 alkyl nitrates (Flocke et al., 1998; Simpson et al., 2006; Archibald et al., 2007). Archibald et al. (2007) reported a model run which suggested that the formation of methyl nitrate via the two production routes would be equal at NOx values of 37 ppb. It is unclear from the paper whether they mean the NOx mixing ratio or NO2 mixing ratio but either would suggest a value of a similar magnitude to that shown in Table 6.6.]

\subsubsection*{Loss reactions}
Being a reservoir for the short lived NOx and having atmospheric lifetimes ranging from several days to about a month \citep{Reeves2007} alkyl nitrates can be transported over long distances, and then lost through photolysis and reaction with $OH$:
\begin{equation} \label{reac:RONO2+hv=RO+NO2}
RONO_2 + hv \rightarrow RO + NO_2
\end{equation}
\begin{equation} \label{reac:RONO2+OH=RCHO+NO2}
RONO_2 + OH \rightarrow RCHO + NO_2
\end{equation}
It means that alkyl nitrates can ‘release’ $NO_2$ back to the atmosphere after a while, and in association with reactions (10) and (11) start a new $O_3$ formation cycle being quite away from the source.
\citep{Newland2013}
[The main loss processes for the alkyl nitrates are oxidation by the OH radical and photodissociation (e.g. Talukdar et al., 1997a, 1997b). The importance of the OH sink relative to photolysis increases with increasing carbon chain length to become the dominant loss process in C5 alkyl nitrates and above (Roberts and Fajer, 1989).]

The alkyl nitrates are removed from the atmosphere through photolysis and reaction with hydroxyl radical.
The relative importance of these two loss reactions depends generally on the net UV intensity and the size of the carbon chain in alkyl nitrates. Photolysis is relatively more important for nitrates with $< C_4$, while nitrates with longer chains ($\geq C_5$) tend to react faster with $OH$. In case of butyl nitrates both mechanisms are of similar significance \citep{Worton2010}.

@@

\subsection{Motivation}
\citep{Sommariva2008}
[Typically, chemical mechanisms do not contain
enough information about the various formation routes of secondary organic compounds, leading to the underestimation of their concentrations, with possible consequences on the estimates of ozone production in remote regions, or to ineffective ozone-control policies. In this work, a quasi-explicit chemical model based upon the Master Chemical Mechanism (MCM) has been used to study in detail the photochemistry of a number of alkyl nitrates and determine quantitatively their formation processes. This kind of study shows the potential of using mechanisms with a high level of chemical detail to study the formation of pollutants in the atmosphere (Madronich, 2006).]

@@

\citep{Sommariva2008}
[Alkyl nitrates typically constitute only a small
fraction of total NOy <10\%, even though they can be a much larger component in the equatorial marine boundary layer (Blake et al., 2003; Simpson et al., 2006 and references therein). Due to their long lifetime (41 week) alkyl nitrates can transport NOx to remote regions, with potential impacts on regional O3 formation (e.g., Murphy et al., 2006; Simpson et al., 2006)]

\subsubsection*{Oceanic sources}
\citep{Newland2013}
[An oceanic source for C1 – C3 alkyl nitrates was first postulated by Atlas et al. (1993) during the SAGA campaign in which it was observed that elevated mixing ratios of methyl and ethyl nitrates correlated well with equatorial regions of high productivity. This has since been confirmed by field measurements (e.g. Chuck et al., 2002; Blake et al., 2003; Dahl et al., 2005). Chuck et al. (2002) estimate an annual emission flux from equatorial Atlantic and Pacific Ocean regions of 230 and 68 Gg yr-1 for methyl and ethyl nitrate respectively. The oceanic source of more or equal C4 alkyl nitrates is insignificant compared to anthropogenic sources (Blake et al., 2003).
The production mechanism of the alkyl nitrates is thought to be by the reaction ROO + NO in the aqueous phase with the peroxy radical coming from the photolysis of coloured dissolved organic matter (CDOM) and the NO coming from the photolysis of nitrite (Dahl et al., 2003). Direct emission from algae (Chuck et al., 2002) or from bacteria at depth (Dahl et al., 2007) has also been postulated.]

\citep{Atlas1993}
Alkyl Nitrates, Nonmethane Hydrocarbons, and Halocarbon Gases Over the Equatorial Pacific Ocean During Saga 3
February March 1990
The cruise covered a region of the equatorial Pacific Ocean from 15N to 10S latitude and 144 to 165W longitude. On this cruise we collected samples for the measurement of alkyl nitrates (RONO2) (C2-C5), nonmethane hydrocarbons (NMHC) and several halocarbon gases.
Results:
- Alkyl nitrates were found at levels higher than predicted from steady state calculations based on measured mixing ratios of hydrocarbons and NO. The measured levels of RONO2 suggest long-range transport as one mechanism contributing to elevated concentrations of alkyl nitrates in the remote troposphere. However, the distributions of C2 and C3 alkyl nitrates over the equator were  similar to the organobromine gases. This distribution suggests
a possible oceanic source for alkyl nitrates to the atmosphere.
Model: 1D 41 constituents incl. 3 ANs

\citep{Chuck2002} September, October
Measurements of methyl and ethyl nitrate in seawater and air samples along two Atlantic Ocean transects provide the FIRST direct evidence for an oceanic source of these compounds. A simple box model calculation suggests that the equatorial source could be a major component of the local atmospheric alkyl nitrate budget.
In the remote atmosphere, the concentration of methyl nitrate (MeONO2) is expected to be low because of low concentrations of NOx and a branching ratio of less than 1\% for the formation of MeONO2
from the reaction of alkylperoxy radicals with NO (8). However, unexpect- edly high levels of RONO2 in marine air and observations of a strong positive correlation between ethyl and propyl nitrate and the marine biogenically produced halocarbon, bromoform, in the marine boundary layer over the equatorial Pacific Ocean led Atlas et al. to propose an oceanic source for alkyl nitrates in the atmo- sphere (9). Subsequent atmospheric sampling campaigns, mostly conducted over the remote Pacific Ocean, have provided yet more evidence to support the hypothesis of an oceanic source (2, 10), with the contribution of alkyl nitrates to the NOy
sum reaching 80\% in equatorial and high-latitude regions over the South Pacific (Blake, 1999). Although there have been no pub- lished measurements confirming the oceanic emission of these species, other studies have also invoked a seawater source for methyl, eth- yl, and propyl nitrates (11, 12). The supersaturation of surface waters over large areas of the Atlantic provides the first direct evidence for an oceanic source of MeONO2 and EtONO2. dominant source region is found between the equator and 40S. distinct mechanisms may be respon- sible for the production of methyl and ethyl nitrate, with different water types supporting different mechanisms. Methyl nitrate generally showed profiles more typically associated with algal production, with decreasing concentrations be- low the chlorophyll maximum. In contrast, in waters more strongly affected by continental influences, such as increased loading of dis- solved organic matter, the alkyl nitrates showed maximum concentrations in surface waters, which may suggest that photochemical production is important in shelf and coastal regions.

\citep{Blake2003} oceanic+urban!!!
Observations: concentration distributions of C1-C4 alkyl nitrates observed during the NASA airborne campaigns Pacific Exploratory Mission (PEM) -Tropics A (September– October 1996) and PEM-Tropics B (March–April 1999). The total geographic range for PEM-Tropics A was 45?N–72?S latitude and 153?E–75?W longitude, and for PEM- Tropics B was 40?N–36?S latitude and 149?E–75?W longitude. We observed high methyl nitrate (MeONO2,CH3ONO2) mixing ratios (approximately 50 pptv) at low altitudes in a latitude band between 8?Nto13?S stretching across the equatorial Pacific, illustrating the oceanic source of MeONO2. This source may be associated with the high-nutrient, low-chlorophyll character of equatorial Pacific waters. We discuss MeONO2 and ethyl nitrate (EtONO2,C2H5ONO2), whose abundance is dominated by equatorial oceanic sources, 2-Propyl nitrate (2-PrONO2,2-C3H7ONO2), which has significant oceanic and northern hemispheric (NH) sources associated with urban/ industrial hydrocarbon emissions, and 2-butyl nitrate (2-BuONO2 2-C4H8ONO2), which has mostly NH (northern hemisphere) sources.
the ocean as the major source for MeONO2
Ethyl nitrate has a dominant oceanic source in equatorial regions, but at mid NH latitudes a small additional source is associated with urban/industrial emissions.

\subsubsection*{Biomass burning source}
\citep{Newland2013}
[Alkyl nitrates also have a biomass burning source (Friedli et al., 2001; Simpson et al., 2002). Simpson et al. (2002) report emissions of C1 – C4 alkyl nitrates from savanna fires in Australia, with methyl nitrate being primarily emitted during the ’flaming’ stage and ethyl – butyl nitrates mainly emitted during the ‘smouldering’ stage. They estimate global emissions from this source to be on the order of 8 Gg yr-1]

\subsubsection*{Lifetime}
\citep{Newland2013}
[Species within this group provide a transport reservoir for the short lived NOx compounds due to their relatively longer lifetimes. This allows NOx to be transported from polluted areas to remote background sites. Alkyl nitrates have been reported as making up only a few percent of the NOy ('odd' nitrogen) budget in urban and rural areas (Shepson et al., 1993; Thornberry et al., 2001), however due to their relatively long lifetimes and their continued production during transport of polluted air masses, as alkanes are converted to alkyl nitrates (e.g. Bertman et al., 1995), they can comprise up to tens of percent of the NOy budget in polar regions and in the remote marine troposphere (Talbot et al., 2000). Therefore alkyl nitrates could potentially be a substantial source of NOx in remote background regions contributing to tropospheric ozone formation in these regions.]
\citep{Newland2013}
[The lifetimes of the alkyl nitrates vary with latitude and season due to the seasonality of the OH sink and photolysis. Clemitshaw et al. (1997) report a range of lifetimes with respect to OH and to photolysis. Photolysis lifetimes vary from about ten days for butyl and pentyl nitrate in the summer to 1 – 2 years in the winter at 65N with lifetimes also decreasing with altitude. The lifetimes due to reaction with OH (tauOH) at 65N range from tens of years in
winter to about 10 days in the summer. tauOH increases with altitude as the OH concentration decreases. At the equator tauOH is similar all year round, on the order of a few days.]
\citep{Chuck2002}
Use of the mean sea-to-air fluxes for 0 to 40S (Table 1) and measured atmospheric concentrations results in a lifetime of be- tween 4.5 and 25 days for MeONO2 and between 5 and 10 days for EtONO2 (22). Literature estimates of the atmospheric life- time of MeONO2, based on known atmospheric destruction processes, are poorly con- strained, ranging from 6 to 29 days (23, 24). For EtONO2
, the atmospheric lifetime with respect to photodissociation at the equator and at 40S has been estimated to be 7 and 12 days, respectively, for October (25).

\subsubsection*{Global distribution}
\citep{Newland2013}
[There have been a number of studies of the latitudinal and seasonal variations of alkyl nitrates in marine air (Blake et al., 2003), in the Arctic (Swanson et al., 2003) and in the Antarctic (Beyersdorf et al., 2010).
Blake et al. (2003) report extensive measurements of C1 – C4 alkyl nitrates from the Pacific between 45N and 72S during the PEM-Tropics A and B projects. This study observed an equatorial band between about 8N and 13S in which the oceans are a source of alkyl nitrates. The highest mean methyl and ethyl nitrate measurements were found in this region. The highest mixing ratios of 2-propyl nitrate were found in the most northerly regions of the study (i.e. 40-45N) but it also showed significantly elevated mixing ratios in the same equatorial band as the C1 – C2 nitrates. The high mixing ratios in the northern mid-latitudes are explained by production from anthropogenic polluted outflow. 2-butyl nitrate showed almost no enhancement in equatorial regions and again had the highest mixing ratios in the most northerly regions of the study.
Swanson et al. (2003) and Beyersdorf et al. (2010) observe a strong seasonality of alkyl nitrate mixing ratios in the Arctic and Antarctic respectively. Polar alkyl nitrate mixing ratios show a similar seasonality to the alkanes, with a spring-time peak just before polar sunrise and lows in the summer time. Winter time mixing ratios of 2-propyl and 2-butyl nitrate are about 10 ppt, while methyl and ethyl nitrate are about 5 – 6 ppt. In the summer 2-propyl and 2-butyl nitrate mixing ratios fall close to zero, while methyl and ethyl nitrate mixing ratios fall to about 3 ppt. The amplitude of the seasonal cycle reflects the reaction rate with OH, i.e. it increases with increasing carbon number.]

\subsection{Mean AN concentrations}
mean ANs concentrations: \citep{Reeves2007}, \citep{Roberts1998}
\citep{Atherton1989}
Alkyl nitrate concentrations of 0.8-37 pptv (2-butyl and 2- and 3-pentyl nitrates) in remote marine environments [Atlas, 1988] and approximately 2 ppbv in urban environments [Hanst et al., 1982] have been measured.
\citep{Chuck2002}
The atmospheric mole fraction of MeONO2 ranged from 1.2 to 43 pptv (Fig. 2,Aand B) and 0.8 to 10 pptv for EtONO2 (Fig. 2, D and E). These values lie within the range of concentrations reported in the literature for MeONO2 (0.2 to 50 pptv) and EtONO2 (0.3 to 12.2 pptv) in the remote marine troposphere (2, 8, 9).
\citep{Blake2003}
Pacific Ocean methyl nitrate 50 pptv

\subsubsection{Comparison methodology}
\citep{Sommariva2008}
[The ratios between the alkyl nitrates and their
main alkane precursors can also be plotted against one another, for example against 2-butylnitrate/ n-butane as in Fig. 4. In this way the modelled and the measured concentrations can be compared without taking into account their variation with time. A similar approach was used in previous studies to investigate the sources of alkyl nitrates and the photochemical processing of polluted air masses (e.g., Bertman et al., 1995; Flocke et al., 1998; Roberts et al., 1998; Stroud et al., 2001; Simpson et al., 2003; Reeves et al., 2007). These studies compared plots of the measured ratios against 2-butylnitrate/n-butane to the ratios calcu- lated with a simple sequential reaction model which assumed a single precursors for each alkyl nitrate. Usually, the underestimation of the ratios for short- chain compounds (propylnitrates/propane, ethylni- trate/ethane) was interpreted as an indication of photochemical sources other than the main alkane precursor.]

\subsubsection{Modelling studies}
A huge number of modelling studies has been done about alkyl nitrates.

\citep{Atherton1989}
The transport from urban areas can lead to alkyl nitrate concentrations of 4-83 pptv after 120- 240 hours. These levels are far higher than those possible from marine in-situ production. Additionally, alkyl nitrates may be more abundant than PAN in remote marine environments.
Model: box
Chemistry: explicit chemical reaction mechanism
[Carter et al., 1986. A variety of alkyl nitrates, including methyl-, butyl-, pentyl-, hexyl-, heptyl-, octyl-, and nonyl-nitrates are allowed to form from fourteen different alkyl peroxy radicals. + ocean emission sources fluxes of c
for ethene, ethane, propene, and propane, NO, NO2
Initial conditions: clear sky conditions, a quantum yield of 1, solar zenith angle of 45 degrees, and total ozone column of 8*10 cm-2. Daytime+nigthtime, invertion height, dilution. 298 K. No deposition.
Scenarios: Two, MLOW and MHIGH, represent the marine in-situ production of alkyl nitrates. The total hydrocarbon concentration was 17.6 ppbC. MLOW had NO = 4 pptv and NO2 = 4.6 pptv based on measurements over the Pacific Ocean. MHIGH had NO = 46 pptv, NO2 = 54 pptv, and PAN = 50 pptv. The two other scenarios, UHIGH and ULOW, simulate the transport of species from urban to remote environments. UHIGH contained 2150 ppbC and 140 ppbv NOx, and ULOW contained 270 ppbC and 18 ppbv NOx, giving each a HC/NOx ratio of 15.
Simulation time: 12 hours MLOW and MHIGH, 240 hours UHIGH and ULOW.

Results:
- The total alkyl nitrate concentration present after 12 hours is less than 0.4 pptv for MHIGH and less than 0.1 pptv for MLOW. These levels are much lower than observations, indicating that marine in-situ production can account for only a fraction of the total marine alkyl nitrate concentration.
- IHIGH The total alkyl nitrate concentration peaks during the first 6 hours and then drops to 3.1 ppbv after 12 hours. This value is com- parable with the 2 ppbv alkyl nitrate concentration de- termined in Los Angeles [Hanst et al., 1982]
- The model predicts alkyl nitrate concentrations of 21-83 pptv for UHIGH and 4-12 pptv for ULOW for the hours 120-240.
- General: The model predictions of 12-83 pptv (af- ter 120 hours) and 4-21 pptv (after 240 hours) are of the same order as observations.

\citep{Sommariva2008}
Validation: typical urban plume in the North-East of the United States, measurements taken during three flights of the NOAA WP-3D aircraft, which sampled a plume from the New York City area during the New England Air Quality Study (NEAQS) 2004 campaign
- Seven alkyl nitrates (ethyl- nitrate, i-propylnitrate, n-propylnitrate, 2-butylni- trate, 2-pentylnitrate, 3-pentylnitrate, 3-methyl-2- butylnitrate) were measured during these flights together with a full suite of VOCs and other chemical parameters, such as CO, O3,NOx, HNO3 and PANs (Neuman et al., 2006).
- The model results have been qualitatively compared to measurements of alkyl nitrates taken on 20–22 July 2004, when the NOAA WP-3D aircraft made three flights aimed to inter- cept the New York City plume.
Model: subset! of MCM v3.1 (The MCM is one of the most detailed chemical mechanism for tropospheric chemistry available)
- to study the formation processes and photochemical evolution of alkyl nitrates
Chemistry: oxidation mechanisms of 64 VOCs (50 hydrocar-
bons and 14 oxygenates) plus CH4 and CO, The inorganic chemistry mechanism was taken from the IUPAC recommendation (Atkinson et al., 2003). A dimethyl sulphide (DMS) mechanism (Sommariva et al., 2004) was also included. The photolysis rates were calculated using the MCM parametrization (Saunders et al., 2003) for end of July 2004 at the latitude and longitude of Boston, MA. only gas-phase!!! Measurements made during NEAQS 2004 showed that there was very little removal of, for example,HNO3 by dry deposition in plumes decoupled from the marine boundary layer during transport from New York City (NYC) to the North Atlantic (Neuman et al., 2006). Therefore, removal by dry deposition was neglected in the model.
Initialisation: measurements taken in the Boston metropolitan area during the 2002 and 2004 NEAQS campaigns. The initial concentrations of VOCs were determined using emission ratios; CH4 and O3 were set to typical values for urban conditions (1800 and 90 ppb) (Table 1). The model has been initialized using measurements and VOCs emission ratios from the 2002 and 2004 New England Air Quality Studies.
Simulation time: 5 days.
Scenarios: 2: A (higher OH, good with obs, reference), B
Time: 20-22 July.

Results:
- While long-chain (C5) alkyl nitrates are produced for 90\% or more from the oxidation of a single parent alkane, short-chain <C4 alkyl nitrates can be formed from several precursors. The relative importance of each production route has been quantitatively determined thanks to the high level of chemical detail provided by the MCM. These secondary routes
to the formation of alkyl nitrates include the oxidation of longer-chain alkanes and oxygenated intermediates, like carbonyls, peroxides and carboxylic acids.
- The best agreement is in the case of 3-methyl-2-butylnitrate and the discrepancy between the modelled and the measured ratios clearly increases with decreasing carbon number .
- RH/AN, Except in the case of ethylnitrate, the model can reproduce the measured rate of increase in the alkyl nitrate to alkane ratios fairly well, which suggests that the kinetic parameters used in the MCMmodel are overall correct.
- By considering together Figs. 3 and 4 it can be reasonably concluded that the chemistry of alkyl nitrates is well represented in the MCM model.
- Using the explicit description of the chemistry provided by the MCM, it has been possible to determine quantitatively the photochemical formation and the evolution of alkyl nitrates and the relative importance of their different formation pathways. The results are summarized in Table 2. Large alkyl nitrates are produced almost exclusively from the reaction of a single parent
alkane. Smaller <=C4 alkyl nitrates also have other significant precursors, like longer-chain alkanes and alkenes. For n-propylnitrate and ethylnitrate, pro- panal and butanal, which are largely emitted, are important sources within the first 24 h. For all alkyl nitrates, the role of organic peroxides and car- boxylic acids formed by the reactions of peroxy radicals with HO2 at low [NOx] increases with time. These species contribute to sustain the production of alkyl nitrates far from the emission points of the parent VOCs. The relative importance of the different pathways
and how they change with time are dependent on the initial conditions of the model. In a scenario with lower O3 and CO and higher NOx the
importance of the direct routes (alkane ?OH) is typically lower during the first two–three days, and the contributions of peroxides and carboxylic acids become significant later and are on average less important. These results suggest that secondary chemistry in the formation of the alkyl nitrates, particularly oC4, cannot be neglected and should be included in the chemical models, due to its potential impact on ozone-controlling strategies.

\citep{Newland2013}
Model1: 2-D atmospheric model
- to study the formation, loss and transport of the alkyl nitrates to the Arctic
- to derive northern hemisphere emissions of the six alkanes.
Chemistry:
- (with fixed NOx and OH field taken from full chemistry run with model by Hough and Johnson (1990))
Validation of ANs: comparison with measurements of alkyl nitrates in the Arctic.
Performance: recreate both the measured seasonality of the alkyl nitrate, and the peak mixing ratio, reasonably well.
Simulation time: 6 days?

Model2: box
Chemistry:
- alkanesmethane, ethane, propane, butane and pentane and ANs explicitly described (RO+NO2 too)

Model3: firn model
- to derive atmospheric histories of six alkanes and six alkyl nitrates from measurements in firn air from the NEEM site, Greenland (methyl, ethyl, 2-propyl, 2-butyl, 2+3-pentyl and 3- methyl-2-butyl nitrate), 1950-2008.

Results:
- There are two ways in which NOx mixing ratios can limit alkyl nitrate production: either through air masses being NO limited with respect to the formation of a given alkyl nitrate or by the NOx in an air mass being 'used up' (track back to Reeves)
-- Alkyl nitrate production becomes NOx saturated above a given NO mixing ratio. This limit increases with increasing carbon number.
-- The time taken for an air mass to 'use up' all the initial NOx is dependent mainly on the initial NOx mixing ratio.
- At high [NO2] (> 27 ppb), methyl nitrate production is dominated by the reaction of the alkoxy radical with NO2
- The regions that contribute most to modeled alkane and alkyl nitrate mixing ratios in the Arctic are Europe and Russia
- The importance of the RO + NO2 pathway compared to RO2 + NO is entirely dependent on the branching ratio (alpha3) of the RO2 + NO reaction pathway. The smaller alpha3 is, the more important the RO + NO2 pathway is.
-- At 10 ppb of NO2 the RO + NO2 pathway accounts for 30 – 60\% of methyl nitrate production. At 100 ppb it accounts for 85 – 95\% of methyl nitrate production and almost 20\% of ethyl nitrate production.
- The rate of alkyl nitrate production in a polluted air mass is proportional to the OH concentration in that air mass.

\subsubsection*{Other}

@@

They are produced photochemically through the oxidation of their parent hydrocarbons \citep{Roberts1990} or emitted directly from equatorial oceans \citep{Blake2003} and biomass burning \citep{Simpson2002}. Being involved in the ozone production chemistry, these species can act as reservoirs of nitrogen oxides which availability largely affects the production of ozone \citep{Reeves2007}. Since the alkyl nitrates have atmospheric lifetimes ranging from several days to about a month, they can redistribute nitrogen oxides presented over polluted continental regions to clean remote marine environments, and therefore are considered as one of the tracers for determining the anthropogenic influence \citep{Atherton1989, Reeves2007, Worton2005}.
